Local Spectral Clustering of Density Upper Level Sets

### 2020-10-31 jmlr/revision/local_spectral_density_clustering_jmlr_revision.tex
Separate folder for JMLR revision.

\documentclass[11pt,twoside]{article}
\usepackage{fancyhdr}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue,bookmarks=false]{hyperref}
\usepackage{amsfonts,epsfig,graphicx}
\usepackage{afterpage}
\usepackage{amsmath,amssymb,amsthm} 
\usepackage{fullpage}
\usepackage{epsf} 
\usepackage{graphics} 
\usepackage{amsfonts,amsmath}
\usepackage[sort,numbers]{natbib} 
\usepackage{psfrag,xspace}
\usepackage{color,etoolbox}

@@

\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

@@

@@

\begin{center} {\Large{\bf{Local Spectral Clustering of Density Upper Level Sets}}}

\vspace*{.3cm}

{\large{
\begin{center}
Alden Green~~~~~ Sivaraman Balakrishnan~~~~~ Ryan J. Tibshirani\\
\vspace{.2cm}
\end{center}

\begin{tabular}{c}
Department of Statistics and Data Science \\
Carnegie Mellon University
\end{tabular}

\vspace*{.2in}

\begin{tabular}{c}
\texttt{\{ajgreen,sbalakri,ryantibs\}@stat.cmu.edu}
\end{tabular}
}}

\vspace*{.2in}

\today
\vspace*{.2in}

\begin{abstract}
We analyze the Personalized PageRank (PPR) algorithm, a local spectral method
for clustering, which extracts clusters using locally-biased random walks around
a given seed node.  In contrast to previous work, we adopt a classical
statistical learning setup, where we obtain samples from an unknown
distribution, and aim to identify connected regions of high-density (density
clusters).  We prove that PPR, run on a neighborhood graph, extracts
sufficiently salient density clusters, that satisfy a set of natural geometric
conditions. We also show a converse result, that PPR can fail to recover
geometrically poorly-conditioned density clusters, even asymptotically. Finally,
we provide empirical support for our theory.
\end{abstract}
\end{center}

@@

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.495\textwidth]{lower_bound_1}
  \includegraphics[width=0.495\textwidth]{lower_bound_2}
  \caption{\it\small The density $f$ in \eqref{eqn:lb_density}, for
    $\rho=1$, and two different choices of $\epsilon$ and $\sigma$. Left:
    $\epsilon = 0.3$ and $\sigma = 0.1$; right: $\epsilon = 0.2$ and 
    $\sigma = 0.2$.} 
  \label{fig: hard_case}
\end{figure}

@@

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.32\textwidth]{sample2}
  \includegraphics[width=0.32\textwidth]{sample1}
  \includegraphics[width=0.32\textwidth]{sigma_normalized_cut_plot}
  \includegraphics[width=0.32\textwidth]{sigma_mixing_time_plot}
  \includegraphics[width=0.32\textwidth]{diameter_normalized_cut_plot}
  \includegraphics[width=0.32\textwidth]{diameter_mixing_time_plot}
  \caption{\it\small Top left and top middle: samples from a geometrically
    well- and poor-conditioned cluster. The points in $\Cset$ are colored in 
    red, points in $\Csig \setminus \Cset$ are colored in yellow, and the
    remaining points in blue. Other panels: empirical normalized cut and mixing
    time, as a function of $\sigma$ or $\rho$, versus their theoretical upper
    bounds.} 
  \label{fig:bounds}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.24\textwidth,scale = .5]{row1_true_density_cluster}
  \includegraphics[width=0.24\textwidth]{row1_ppr_cluster}
  \includegraphics[width=0.24\textwidth]{row1_conductance_cluster}
  \includegraphics[width=0.24\textwidth]{row1_density_cluster}
  \includegraphics[width=0.24\textwidth]{row2_true_density_cluster}
  \includegraphics[width=0.24\textwidth]{row2_ppr_cluster}
  \includegraphics[width=0.24\textwidth]{row2_conductance_cluster}
  \includegraphics[width=0.24\textwidth]{row2_density_cluster}
  \includegraphics[width=0.24\textwidth]{row3_true_density_cluster}
  \includegraphics[width=0.24\textwidth]{row3_ppr_cluster}
  \includegraphics[width=0.24\textwidth]{row3_conductance_cluster}
  \includegraphics[width=0.24\textwidth]{row3_density_cluster}
  \caption{\it\small True density (column 1), PPR (column 2), normalized
    cut (column 3) and estimated density (column 4) clusters for 3 different 
    simulated data sets. Seed node for PPR denoted by a black cross.} 
  \label{fig:moons}
\end{figure}

@@

\bibliographystyle{plainnat}
\bibliography{local_spectral_bibliography}

@@

	\noindent Assuming $\Csig$ satisfies~\ref{asmp: bounded_density}, with probability at least $1 - 2n\exp\set{-\frac{2 \delta^2 \lambda_{\sigma} \nu_d r^d n}{75(1 + \frac{\delta}{3})}}$,
	\begin{equation}
	\label{eqn:degwt_bound}
	\frac{6}{25}(1 - \delta)\lambda_{\sigma}\nu_dr^dn \leq \degminwt \leq d_{\max} \leq (1 + \delta) \Lambda_{\sigma} \nu_dr^dn.
	\end{equation}
	Assuming $\Cset'$ satisfies~\ref{asmp: C'_bounded_density}, with probability at least $1 - n\exp\set{-\frac{\delta^2 \lambda_{\sigma} \nu_d r^d n}{3(1 + \frac{\delta}{3})}}$,
	\begin{equation}
	\label{eqn:degpr_bound}
	\degminpr \geq (1 - \delta)\lambda_{\sigma}\nu_dr^dn.
	\end{equation}
\end{lemma}
\begin{proof}{(of Lemma~\ref{lem:ball_bounds_in_probability})}
	\item[]
	\paragraph{Proof of \eqref{eqn:degwt_bound}, \eqref{eqn:degpr_bound}:}
	Under~\ref{asmp: bounded_density}, we have that for every $x_j \in \Csig[\Xbf]$
	\begin{equation*}
	\mathbb{E}(\widetilde{\deg}_{n,r}(x_j)) \leq \Lambda_{\sigma} \nu_d r^d;
	\end{equation*}
	furthermore, recalling the weighted local conductance $\ell_{\Pbb,r}(x) = \Pbb(B(x,r) \cap \Csig)$, we have,
	\begin{equation*}
	\mathbb{E}(\widetilde{\deg}_{n,r}(x_j)) = \ell_{\Pbb,r}(x_j) \geq \frac{6}{25} \lambda_{\sigma} \nu_d r^d
	\end{equation*}
	where the last inequality follows from Lemma~\ref{lem: local_conductance}.  Under~\ref{asmp: C'_bounded_density} we have for every $x_i \in \Cset'[\Xbf]$
	\begin{equation*}
	\mathbb{E}(\widetilde{\deg}_{n,r}(x_i)) \geq \lambda_{
		\sigma} \nu_d r^d.
	\end{equation*} 
	Then \eqref{eqn:degwt_bound} and \eqref{eqn:degpr_bound} each follow from Lemma~\ref{lem:bernstein_union}.

@@

\begin{figure}
	\centering
	\begin{adjustbox}{minipage=\linewidth,scale=0.8}
		\begin{subfigure}{.24\linewidth}
			\includegraphics[width=\linewidth]{row4_true_density_cluster}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.24\linewidth}
			\includegraphics[width=\linewidth]{row4_ppr_cluster}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.24\linewidth}
			\includegraphics[width=\linewidth]{row4_conductance_cluster}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.24\linewidth}
			\includegraphics[width=\linewidth]{row4_density_cluster}
			\caption{}
		\end{subfigure}
		\caption{}
		\label{fig:moons_hd}
	\end{adjustbox}
\end{figure}

>>>>

\documentclass[11pt,twoside]{article}
\usepackage[preprint]{jmlr2e}
\usepackage{color}
\usepackage{fancyhdr}
\usepackage{amsfonts,epsfig,graphicx}
\usepackage{afterpage}
\usepackage{amsmath,amssymb} 
% \usepackage{fullpage}
\usepackage{epsf} 
\usepackage{graphics} 
\usepackage{amsfonts,amsmath}
% \usepackage[sort,numbers]{natbib} 
\usepackage{psfrag,xspace}
\usepackage{color,etoolbox}

@@

%\newtheorem{theorem}{Theorem} 
%\newtheorem{lemma}{Lemma}
%\newtheorem{proposition}{Proposition}
%\newtheorem{corollary}{Corollary}
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}
%\newtheorem{remark}{Remark}

@@

\ShortHeadings{Local Spectral Clustering of Density Level Sets}{Green, Balakrishnan and Tibshirani}
\firstpageno{1}

@@

\title{Local Spectral Clustering of Density Upper Level Sets}

\author{\name Alden Green \email ajgreen@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213
	\AND
	\name Sivaraman Balakrishnan \email sbalakri@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213
	\AND
	Ryan J. Tibshirani \email ryantibs@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213}

\maketitle

\begin{abstract}
We analyze the Personalized PageRank (PPR) algorithm, a local spectral method
for clustering, which extracts clusters using locally-biased random walks around
a given seed node.  In contrast to previous work, we adopt a classical
statistical learning setup, where we obtain samples from an unknown
distribution, and aim to identify connected regions of high-density (density
clusters).  We prove that PPR, run on a neighborhood graph, extracts
sufficiently salient density clusters, that satisfy a set of natural geometric
conditions. We also show a converse result, that PPR can fail to recover
geometrically poorly-conditioned density clusters, even asymptotically. Finally,
we provide empirical support for our theory.
\end{abstract}

\begin{keywords}
	graphs, spectral clustering, density clustering, Personalized PageRank, unsupervised learning
\end{keywords}

@@

% PUT FIGURE BACK HERE

@@

% PUT FIGURE BACK HERE

@@

% PUT FIGURE BACK HERE

@@

@@

	\noindent Assuming $\Csig$ satisfies~\ref{asmp: bounded_density}, with probability at least $1 - 2n\exp\set{-\frac{2 \delta^2 \lambda_{\sigma} \nu_d r^d n}{75(1 + \frac{\delta}{3})}}$,
	\begin{equation}
	\label{eqn:degwt_bound}
	\frac{6}{25}(1 - \delta)\lambda_{\sigma}\nu_dr^dn \leq \degminwt \leq d_{\max} \leq (1 + \delta) \Lambda_{\sigma} \nu_dr^dn.
	\end{equation}
	Assuming $\Cset'$ satisfies~\ref{asmp: C'_bounded_density}, with probability at least $1 - n\exp\set{-\frac{\delta^2 \lambda_{\sigma} \nu_d r^d n}{3(1 + \frac{\delta}{3})}}$,
	\begin{equation}
	\label{eqn:degpr_bound}
	\degminpr \geq (1 - \delta)\lambda_{\sigma}\nu_dr^dn.
	\end{equation}
\end{lemma}
\begin{proof}{(of Lemma~\ref{lem:ball_bounds_in_probability})}
	\hfill 
	\paragraph{Proof of \eqref{eqn:degwt_bound}, \eqref{eqn:degpr_bound}:}
	Under~\ref{asmp: bounded_density}, we have that for every $x_j \in \Csig[\Xbf]$
	\begin{equation*}
	\mathbb{E}(\widetilde{\deg}_{n,r}(x_j)) \leq \Lambda_{\sigma} \nu_d r^d;
	\end{equation*}
	furthermore, recalling the weighted local conductance $\ell_{\Pbb,r}(x) = \Pbb(B(x,r) \cap \Csig)$, we have,
	\begin{equation*}
	\mathbb{E}(\widetilde{\deg}_{n,r}(x_j)) = \ell_{\Pbb,r}(x_j) \geq \frac{6}{25} \lambda_{\sigma} \nu_d r^d
	\end{equation*}
	where the last inequality follows from Lemma~\ref{lem: local_conductance}.  Under~\ref{asmp: C'_bounded_density} we have for every $x_i \in \Cset'[\Xbf]$
	\begin{equation*}
	\mathbb{E}(\widetilde{\deg}_{n,r}(x_i)) \geq \lambda_{
		\sigma} \nu_d r^d.
	\end{equation*} 
	Then \eqref{eqn:degwt_bound} and \eqref{eqn:degpr_bound} each follow from Lemma~\ref{lem:bernstein_union}.

@@

% PUT FIGURE BACK HERE

\clearpage

% \bibliographystyle{plainnat}
\bibliography{../../local_spectral_bibliography}
